---
title: "p8130_hw4_rw2844"
author: "Renjie Wei"
date: "10/31/2020"
output: html_document
---


```{r setup, include=FALSE}
library(tidyverse)
library(arsenal)
library(animation)
library(magrittr)
library(patchwork)
library(multcomp)
knitr::opts_chunk$set(
  fig.height = 6,
  fig.width = 8,
  message = F,
  echo = F,
  warning = F
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis",
  digits = 3
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1


**__Proof:__**
$$
\begin{split}
\because by~the~definition,~ y_{ij} - \bar{\bar{y}} = (y_{ij}-\bar{y_i}) + (\bar{y_i} - \bar{\bar{y}})\\
\therefore \sum_j(y_{ij} - \bar{\bar{y}})^2=\sum_j[ (y_{ij}-\bar{y_i}) + (\bar{y_i} - \bar{\bar{y}})]^2 = \sum_j[{ (y_{ij}-\bar{y_i})^2 + (\bar{y_i} - \bar{\bar{y}})^2+2\times (y_{ij}-\bar{y_i})(\bar{y_i} - \bar{\bar{y}})}]\\
\because \sum_j y_{ij}/n_j = \bar{y_i},~and~\sum_j 1= n_j\\
\therefore \sum_j{(y_{ij}-\bar{y_i})(\bar{y_i} - \bar{\bar{y}})}= \sum_j [y_{ij}\times \bar{y_i} - y_{ij}\times \bar{\bar{y}} - \bar{y_i}^2 + \bar{y_i}\times \bar{\bar{y}}]\\
=n_j\times \bar{y_i}^2 -  n_j\times \bar{y_i}\times \bar{\bar{y}} - n_j\times \bar{y_i}^2 + n_j\times \bar{y_i}\times \bar{\bar{y}}\\
= (n_j\times \bar{y_i}^2 -  n_j\times \bar{y_i}^2) + ( n_j\times \bar{y_i}\times \bar{\bar{y}}- n_j\times \bar{y_i}\times \bar{\bar{y}})= 0
\end{split}
$$

Then, we can prove that:

$$ \sum_i\sum_j(y_{ij} - \overline{\bar{y}})^2 = \sum_i\sum_j(y_{ij}-\bar{y_i})^2 + (\bar{y_i} - \overline{\bar{y}})^2$$

# Problem 2



## 1

```{r prob_2_setting}

my_controls <- tableby.control(
               total = F,
               test=F,  
               numeric.stats = c("meansd", "medianq1q3", "range","Nmiss2"),
               cat.stats = c("countpct","Nmiss2"),
               stats.labels = list(
               meansd = "Mean (SD)",
               medianq1q3 = "Median (Q1, Q3)",
               range = "Min - Max",
               Nmiss2 = "Missing",
               countpct = "N (%)"))

```
```{r prob2_plot}
knee_df =
  read.csv("./data/Knee.csv") %>% 
  pivot_longer(
    cols = Below:Above,
    names_to = "physical_status",
    values_to = "rehabitation_days"
  ) %>% 
  janitor::clean_names() %>% 
  mutate(
    physical_status = forcats::fct_relevel(physical_status,"Below","Average","Above"))

tab1 <- tableby( physical_status~ rehabitation_days , data=knee_df, control=my_controls)
sum_tab = summary(tab1, title = "Descriptive Statistics: Knee data (Days)",  text=T)

sum_tab %>% knitr::kable()
```


**Comments:**

From the table, we can see that the Mean of rehabilitation days are different in the three different physical condition groups.

To make it more clear, I made a box plot.

```{r box_plot_descriptive}
knee_df %>% 
  ggplot(aes(x = physical_status, y = rehabitation_days, fill  = physical_status))+
  geom_boxplot(alpha = 0.5)
```


## 2
$$
\begin{split}
H_0: \mu_1=\mu_2=\mu_3,~there~is~no~difference~in~mean~of~the~three~physical~condition~groups\\
H_1: at~least~two~means~are~not~equal\\
\end{split}
$$

```{r prob2_anova}
knee_anova =
  aov(rehabitation_days~physical_status, data = knee_df, alpha = 0.01)
summary(knee_anova)
anova_crit=qf(0.99,2,22)
```
$$
\begin{split}
Between~SS = \sum_{i=1}^k\sum_{j=1}^{n_i}(\bar{y_i} - \bar{\bar{y}})^2=\sum_{i=1}^kn_i\bar{y_i}^2-\frac{y_{..}^2}{n}=(38-31.96)^2+(33-31.96)^2+(23.57-31.96)^2=796\\
Within~SS = \frac{\sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}-\bar{y_i})^2}{n-k}=\sum_{i=1}^k(n_i-1)s_i^2=7\times(5.48)^2+9\times(3.92)^2+6\times(4.20)^2=454\\
Between~Mean~Square = \frac{Between~SS}{k-1}=398\\
Within~Mean~Square = \frac{Within~SS}{n-k}=21 \\
F_{stat} = \frac{Between~Mean~Square}{Within~Mean~Square}=19.3\sim F(k-1,n-k)=F(2,22)\\
F_{crit} = F_{2,22,0.99}= `r anova_crit`
\end{split}
$$

**Decision Rule:**

Reject $H_0$ if $F_{stat}>F_{crit}$

Fail to reject $H_0$ otherwise.



**Conclusion:**

Since $F_{stat}=19.3> F_{crit}=5.791$, at 1% significance level, we reject the null hypothesis and conclude that at least two of mean rehabilitation days from the 3 physical condition groups are different.



## 3

This is the Bonferroni adjust pairwise t-test:

```{r prob2_bonf}
# Bonferroni
bonf =
  pairwise.t.test(
    knee_df$rehabitation_days,
    knee_df$physical_status,
    p.adjust.method = "bonferroni",
    conf.level = 0.99
    )

show(bonf)
```


This is the Tukey's test:
```{r prob2_tukey}
##Tukey's
tukey =
  TukeyHSD(knee_anova)

show(tukey)
plot(tukey)
```

This is the Dunnett's test:
```{r prob2_dunnett}
##Dunnet
summary(glht(knee_anova), linfct = mcp(Group = "Dunnett"))
```

**Comments:**

* **Similarities:** All these three test are multiple comparisons adjustment methods. They are used to find out which two groups are different in mean.

* **Differences:** Tukey's method is less conservative than Bonferroni, which means it has a less p-values. And Dunnett's method need to define a control arm before the test.


## 4

The total mean of the days of rehabilitation is 31.96 days. And by using an ANOVA, we found that the rehabilitation days are not all the same between three different physical condition groups at 1% significance level, whereas whom with below average physical status takes 38 days, whom with average takes 33 days compared with 23 days in above average group. After multiple comparisons, we confirm the conclusion and we reject that the patients begin with different physical conditions need the same recover time.


# Problem 3

## 1

We are going to compare the proportions between 3 different groups---- Major swelling, Minor swelling and No swelling. We should create a $R\times C$ Contingency table and conduct a Chi-Squared test of homogeneity.

## 2
```{r}
swell_df =
  matrix(
    c(54,16,42,32,134,142),ncol = 3,
    dimnames = list(
      c("Vaccine","Placebo"),
      c("Major","Minor","No")
         )
    )
## Table for Observed values
swell_df %>% 
  as_tibble() %>% 
  mutate(Groups =c("Vaccine","Placebo")) %>% 
  relocate(Groups) %>%
  knitr::kable(caption = "Observed Values")
chisq_swell =
  chisq.test(swell_df)

## Table for Expect values
chisq_swell$expected %>% 
  as_tibble() %>% 
  mutate(Groups =c("Vaccine","Placebo")) %>% 
  relocate(Groups) %>% 
  knitr::kable(caption = "Expected Values")

p_chi =chisq_swell$p.value
```

## 3

Now we are going to use Chi-Square test:

$$
\begin{split}
H_0:p_{1j} =p_{2j}=...=p_{5j}=p_{.j},i=1,2,~j=1,2,3, the~proportions~of~different~swelling~conditions~are~the~same~among~vaccine~and~placebo~groups\\
H_1: p_{ij}\neq p_{i'j},~j=1,2,3,i\neq i', ~not~all~vote~rates~are~equal\\
\chi^2_{stat} = \sum_i^{R}\sum_j^{C}\frac{(O_{ij}-E_{ij})^2}{E_{ij}} \sim \chi^2_{ (R-1)\times(C-1)},where~df = (R-1)\times(C-1)=1\times 2 = 2\\
\chi^2_{stat}=(54-38.3)^2/38.3+(42-40.5)^2/40.5+(134-151)^2/151+(16-31.7)^2/31.7+(32-33.5)^2/33.5+(142-125)^2/125=`r chisq_swell$statistic`\\
\chi^2_{crit}=\chi^2_{ (R-1)\times(C-1),1-\alpha}=\chi^2_{ 2,0.95}=`r qchisq(0.95, 2)`
\end{split}
$$

**Decision Rule:**

Reject$H_0$ if $\chi^2_{stat}>\chi^2_{crit}$.

Fail to reject $H_0$ otherwise.

And the p-value is:
$$
\begin{split}
p-value
= \int_{x^2}^\infty2*{Z^2}=\int_{x^2}^\infty{\frac{1}{\pi}e^{-s^2}}ds\\
=e^{-x^2/2}=`r p_chi`
\end{split}
$$

**Conclusion:**

Since in our situation $\chi^2_{stat}>\chi^2_{crit}$, and the p-value is $`r p_chi`$we reject the null hypothesis at 5% significance level, and conclude that not all the proportions of different swelling conditions are the same among vaccine and placebo group. Or the swelling condition is significantly different between vaccine and placebo group.



